---
title: "P8105_hw1_cz2544"
author: "Triciazhai"
date: "9/13/2018"
output: html_document
---


## Problem_1 

###1.1 Create a data frame comprised of:

* A random sample of size 10 from a uniform[0, 5] distribution  
* A logical vector indicating whether elements of the sample are greater than 2  
* A (length-10) character vector   
* A (length-10) factor vector   

```{r hw_1_problem_1.1}
library(tidyverse)
hw1_df1 = tibble(
  vec_samp = runif(10, min=0, max=5),
  vec_logi = vec_samp > 2,
  vec_char = head(month.abb, 10),
  vec_fact = factor(c("cold", "cold", "cool", "warm", "warm", "hot", "hot", "hot", "cool", "cool"))
)
hw1_df1
```

###1.2 Try to take the mean of each variable in the dataframe

```{r hw_1_problem_1.2}
mean_vec = c(mean(hw1_df1$vec_samp), mean(hw1_df1$vec_logi), mean(hw1_df1$vec_char), mean(hw1_df1$vec_fact))
mean_vec
```

**Comment:**   
Mean value could be achieved for numeric/logical vectors and date, date-time and time interval objects, but not for charactor and factor vectors. For logical vectors, true = 1, and false = 0.

###1.3 Write a code chunk that applies the as.numeric function to the logical, character, and factor variables

```{r hw_1_problem1.3, results='hide'}
as.numeric(hw1_df1$vec_logi)
as.numeric(hw1_df1$vec_char)
as.numeric(hw1_df1$vec_fact)
```

###1.4 Convert your character variable from character to factor to numeric

```{r hw_1_problem1.4}
as.factor(hw1_df1$vec_char)
as.numeric(hw1_df1$vec_char)
```

**Comment:**   
When characters are converted to factors, r recognized every different character as a factor level.
Characters cannot be converted to numbers, so 10 "NA"" is the result.

###1.5 Convert your factor variable from factor to character to numeric

```{r hw_1_problem1.5}
as.character(hw1_df1$vec_fact)
as.numeric(hw1_df1$vec_fact)
```

**Comment:**   
When factors are converted to charactors, the will be process as words in "", no longer as different levels.
Factor vecters can be converted to numeric vectoers, as r recognize different factor levels of n, numbers from 1 to n has been asigned to different levels of factors. When convert factor vectors to numeric vectors, the asigned numbers for facter will be the new value.

## Problem_2 

###2.1 Create a data frame comprised of:

* x: a random sample of size 1000 from a standard Normal distribution   
* y: a random sample of size 1000 from a standard Normal distribution   
* A logical vector indicating whether the x + y > 0   
* A numeric vector created by coercing the above logical vector   
* A factor vector created by coercing the above logical vector   

```{r hw_1_problem_2.1}
set.seed(1234)
hw1_df2 = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  xy_logi = x + y > 0,
  xy_nume = as.numeric(xy_logi),
  xy_fact = as.factor(xy_logi)
)
hw1_df2
```

###2.2  Short description of vectors using inline R code, including: 

* the size of the dataset in numbers of rows and columns is `r dim(hw1_df2)`,
* the mean and median of x are `r mean(hw1_df2$x)` and `r median(hw1_df2$x)`
* the proportion of cases for which the logical vector is TRUE is `r sum(hw1_df2$xy_logi)/length(hw1_df2$xy_logi)`

###2.3 Make a scatterplot of y vs x:

color points using the logical, numeric and factor variables.

```{r hw_1_problem_2.3}
ggplot(hw1_df2, aes(x, y, color = xy_logi)) + geom_point() 
ggplot(hw1_df2, aes(x, y, color = xy_nume)) + geom_point() 
ggplot(hw1_df2, aes(x, y, color = xy_fact)) + geom_point()
```

**Comment:**    
The color scale for logical and facter vectors are discrete, one color for each value.
The color scale for numerical vector is a two color gradient. Although the numeric vector in this case has only two values, it is processed as a contineous variable by R.

###2.4 Export first scatterplot

```{r hw_1_problem_2.4}
ggplot(hw1_df2, aes(x, y, color = xy_logi)) + geom_point() 
ggsave("hw1_df2_plot_1.pdf", path = NULL)
```

{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

